\chapter{Quotient Types}
\label{qt}




In this chapter, we present a definition of quotient type in an \itt extended with a proof-irrelevant universe of propositions in the sense of \Cref{ittprop}. We prove that given propositional extensionality, all quotients are effective. We also explain the rules of quotient types categorically. A quotient is essentially a coequalizer or given by an adjunction with equality predicate functor \cite{Jacobs94quotientsin}. 
Quotient types in our definition are essentially quotient \emph{sets}, if quotients are not interpreted as sets, for example in \hott. 
We present Voevodsky's impredicative encoding of quotient sets with all essential rules, and also quotient sets defined using higher inductive types. 

\section{Quotients in Type Theory}

\subsection{Rules for quotients}\label{iqs}

The quotient types are defined by the following rules as described in \cite{Jacobs94quotientsin,hof:95:sm}. 


\infrule[Q-\bf{Form}]
{ \Gamma \vdash A  \andalso \Gamma ,x : A , y : A \vdash x \sim y : \bf{Prop} \andalso \sim~\text{is an equivalence}}
{\Gamma \vdash \qset{A}}

Given a type $A$ and with a binary equivalence relation $\sim$ on $A$, we can form the quotient $\qset{A}$. Here we use an infix notation for readability.

The equivalence properties include


\begin{multicols}{2}
\infrule[$\sim$-\bf{refl}]
{ \Gamma \vdash a : A}
{\Gamma \vdash \text{ref}_{\sim}(a) : a \sim a}
\columnbreak
\infrule[$\sim$-\bf{sym}]
{ \Gamma \vdash a : A \andalso \Gamma \vdash b : A \andalso \Gamma \vdash p : a \sim b}
{\Gamma \vdash \text{sym}_{\sim}(p) : b \sim a}
\end{multicols}




\infrule[$\sim$-\bf{trans}]
{ \Gamma \vdash a : A \andalso \Gamma \vdash b : A \andalso \Gamma \vdash c : A \\
\Gamma \vdash p : a \sim b \andalso \Gamma \vdash q : b \sim c}
{\Gamma \vdash \text{trn}_{\sim}(p,q) : a \sim c}

\begin{remark}
Notice that the formation rule is different to Hofmann's version \cite{hof:95:sm} where $\sim$ is not required to be only equivalence relation. In fact his version is just more general which accepts non equivalence relation $R : A \to A \to \Prop$, but $A/R$ has to be understood as the quotient of $A$ by the equivalence closure of $R$.
\end{remark}

\begin{multicols}{2}
\infrule[Q-\bf{Intro}]{\Gamma \vdash a : A}{\Gamma \vdash [ a ] : \qset{A}}
\columnbreak
\infrule[Q-\bf{Ax}]
{\Gamma \vdash a , b :  A  \andalso  \Gamma \vdash  p : a \sim b}
{ \Gamma \vdash \text{Qax}(p) : [a]=_{\qset{A}} [b]}
\end{multicols}


We introduce an ``equivalence class'' for each element of $A$. It is usually denoted as $[ a ]$, or $[ a ]_{\sim}$ for $\sim$ if it is unclear which relation it refers to. 
 Qax states that the ``equivalence classes'' of two terms which are related by $\sim$ are (propositionally) equal.

 Notice that the notation of terms $[a]$ can be confused with the substitution notation like $B[a]$ or $B [[ a ]/x]$. So for a $\Pi$-type $B : (x : A) \to \Set$ and $a : A$, we write $B(a) : \Set$ for $B [[ a ]/x]$ where order of the arguments in brackets matches its definition.

% $for any possibly ambiguous occurrence, we use full notations i.e.\ $[a]_{\sim}$  for terms and $B[a/x]$ for substitution.$


In Hofmann's \cite{hof:95:sm} definition, it comes with an eliminator (also called \emph{lifting}) with its computation rule ($\beta$-rule) and an induction principle (equivalent to a $\eta$-rule): \footnote{We use shorthand notation $\hat{}$ for lifting here} 


\infrule[Q-\bf{elim}]
{\Gamma \vdash  B \andalso \Gamma \vdash f : A \to B \\
\Gamma, a: A, b : A, p : a \sim b \vdash  f^{\sim}(a,b,p) : f(a) =_{B}  f(b) \andalso \Gamma \vdash  q : \qset{A}}
{\Gamma \vdash  \hat{f}(q) : B}

\infrule[Q-\bf{comp}]{\Gamma \vdash  a : A}{\Gamma \vdash  \text{Qcomp}(a)  : \hat{f}([ a ]) \equiv f(a) }


\infrule[Q-\bf{ind}]
{\Gamma,  x : \qset{A} \vdash P : \Prop \andalso \Gamma, a : A \vdash h(a) : P([a]) \andalso \Gamma \vdash  q : \qset{A}}
{\Gamma \vdash \text{Qind}(h,q) : P(q)}

Given a function $f : A \to B$ which respects $\sim$, we can lift it to be a function on $\qset{A}$ as $\hat{f} : \qset{A} \to B$ such that for any element $a : A$, $\hat{f}([ a ])$ computes to the same value as $f(a)$. It allows us to define functions on quotient types by functions on base types (representatives).
%Notice that \emph{function application} is written like "$f ~x ~ y$'' where $x$ and $y$ are two arguments for function $f$. 
Notice that we omit $f^=$ since the computation rule already implies that it is proof-irrelevant.

The induction principle states that for any proposition $P : \qset{A} \to \Prop$. it is enough to just consider cases $ P([ a ])$ for all $a : A$. In other words, $\qset{A}$ only consists of "equivalence classes" i.e.\ $[ a ]$.


Alternatively, a \emph{dependent} eliminator (dependent lifting) serves the same purpose:

\infrule[Q-\bf{dep-elim}]
{\Gamma, x : \qset{A} \vdash B \andalso \Gamma \vdash f : (a : A) \to B([ a ]) \\
\Gamma, a : A, b : A, p : a \sim b \vdash f^= (a,b,p) : f(a)
\stackrel{p}{=}  f(b) \andalso \Gamma \vdash q :
\qset{A}}
{\Gamma \vdash \hat{f}(q) : B(q)}

\infrule[Q-\bf{dep-comp}]{\Gamma \vdash a : A}
{\Gamma \vdash \text{Qdcomp} (a)  : \hat{f}([ a ]) \equiv f(a) }


Notice that $\stackrel{p}{=}$ is an abbreviation for propositional equality which requires substitution in the type of the left hand side by $\text{Qax}(p)$ so that both sides have the same type.
We use the same notation for the two versions of eliminators because they are in fact equivalent.

\begin{proposition}\label{elim-equiv}
The \emph{non-dependent eliminator} with the \emph{induction principle} is equivalent to the \emph{dependent eliminator}.
\end{proposition}
\begin{proof}
1. Assume we have the non-dependent eliminator and the induction principle, $B$ is a dependent type on $\qset{A}$, $f$ is a dependent function of type $(a : A) \to B([ a ])$ and it respects $\sim$ under substitution (i.e.\ $f^=$), $q$ is an element of $\qset{A}$.

Set $B'$ as a dependent product $\Sigma (r : \qset{A}) ~B(r)$,

Then a non-dependent version of $f$ which is of type $A \to B'$ can be defined as

$$f' (a) \defeq [ a ] , f (a)$$

Given $p : a \sim b$, we can conclude that $f'(a) =_{B'} f'(b)$ is inhabited from Qax and $f^=$.

It allows us to lift the non-dependent function $f'$ as $\hat{f'}$ such that 

\begin{equation}\label{f'comp}
\hat{f'} ([ a ]) \equiv [ a ] , f (a)
\end{equation}
Applying first projection on both sides of \ref{f'comp}, the following propositional equality is inhabited:

 $$\pi_1 (\hat{f'}([ a ])) = [ a ]$$

By induction principle, the predicate $P : \qset{A} \to \Prop$ defined as

$$P(q) \defeq \pi_1 ~(\hat{f'}(q)) =_{\qset{A}} q$$

is inhabited for all $q : \qset{A}$.

Finally, to complete the dependent eliminator, we can construct an element of type $B(q)$ by

$$\pi_2 ~(\hat{f'}(q))$$

which has the correct type because $P(q)$ holds. The computation rule is simply derivable from \ref{f'comp}.

2. It is easy to find out that the non-dependent eliminator and induction principle are just special cases of dependent eliminator.

A formalised version of this proof in Agda can be found in \Cref{app:dq}.
\end{proof}


Additionally, a quotient is effective (or exact) if an "equivalence class" only contains terms that are related by $\sim$.

\infrule[Q-\bf{effective}]
{\Gamma \vdash a :  A \andalso \Gamma \vdash b :  A  \andalso p : [a] =_{\qset{A}} [b] }
{\text{eff}(p) : a \sim b}

In fact propositional extensionality implies that all quotients defined with \emph{equivalence} relation are effective.

\begin{theorem}\label{PUEF}
With propositional extensionality, we can prove that all quotient types are effective.
\end{theorem}
\begin{proof}
%Suppose we have a set $A$ with an equivalence relation $\sim : A \to A \to \Prop$, a quotient set is $\qset{A}$.

Suppose we have a quotient type $\qset{A}$, two elements $a, b : A$ and $[ a ] = [ b ]$


Set a predicate $P_a : A \to \Prop$ as 
$$P_a (x) \defeq a \sim x$$

$P_a$ is compatible with $\sim$ since

$x \sim y$

$\Rightarrow a \sim x \iff a \sim y$ (symmetry and transitivity)

$\equiv P_a(x) \iff P_a(y)$ (propositional extensionality)

$\Rightarrow P_a(x) = P_a(y)$

Therefore we can lift $P_a$\footnote{The elimination rule applies to large types} such that for any $x : A$

$$\hat{P} ([ x ]) \equiv a \sim x$$


We can simply deduce $\hat{P} ([ a ]) = \hat{P} ([ b ])$ from assumption $[ a ] = [ b ]$ which by definition
is just

$$a \sim a  = a \sim b$$ 

Finally with eliminator $J$ and $\text{ref}(a) : a \sim a$ we can easily prove

$$a \sim b$$.

\end{proof}

%Alternatively, we can prove it as follows:

%\begin{proof}

%Firstly we prove the equivalence relation is well-defined on the
%quotient types, namely it respects the equivalence relation:

%Suppose we have $a \sim b$ and $c \sim d$, we can deduce $a \sim c \iff
%b \sim d$. Then applying the propositional univalence axiom, we know
%that $a \sim c = b \sim d$, hence the equivalence relation is
%well-defined.

%Because it is well-defined, we can lift it such that

%$[ a ] ~\hat{\sim}~ [ b ] \equiv a \sim b$


%From reflexivity of the equivalence relation, $\forall x : A, x \sim x$, 
%we know that $\forall x : A, [x]~\hat{\sim}~[x]$.

%Assume $[a]=[b]$, using J-eliminator in $[a]~\hat{\sim}~[a]$
%(reflexivity), $[a]~\hat{\sim}~[b]$ which is definitionally equal to $a \sim b$, Hence the quotient is
%effective.
%\end{proof}

Similar to other extensional concepts like functional extensionality, simply adding quotient types into \itt as axioms can also result in \emph{non-canonical construction}.

\begin{theorem}\label{quotientcanonicity}
If we postulate the rules of quotient types, the type theory is not adequate.
\end{theorem}

\begin{proof}
Given a type $A$ and an equivalence relation $\sim$, we postulate $\qset{A}$ exists with all the rules above.

Suppose we have two elements $a, b : A$ such that $p : a \sim b$, we have

$$\text{Qax}(p) : [a] = [b]$$

Define $B : \qset{A} \to \Set$ as

$$B(q)\defeq \N$$

We can observe that $0 : B([a])$, thus by using subst function (see \Cref{subst}), we can obtain a term of $B([b])$:

$$\mathsf{subst}(B,\text{Qax}(p),0) : B([b])$$

which is also a term of $\N$ by definition of $B$.
This term is irreducible to any numeral because $\text{Qax}(p)$ can not be reduced to the canonical term of identity type (refl). Moreover, one can not postulate propositional equality $\text{Qax}(p) = \text{refl}_{[a]}$ or $\text{Qax}(p) = \text{refl}_{[b]}$ because their types are not definitionally equal.
\end{proof}

\section{Quotients are coequalizers}

The rules of quotient types can be characterised in a category-theoretical way.

Categorically speaking, a quotient is a \textbf{coequalizer} in category $\Set$. Let us recall the definition.


\begin{definition}
\textbf{Coequalizer}.
Given two objects $X$ and $Y$ and two parallel morphisms $f, g : \morph{X}{Y}$ , a coequalizer is an object $Q$ with a morphism $q : \morph{Y}{Q}$ such that $q \circ f = q \circ g$ and it is universal: any pair $(Q' , q')$ satisfying $q' \circ f = q' \circ g$ has a unique factorisation $u$ such that $q' = u \circ q$:
\begin{displaymath}
    \xymatrix{X \ar@<0.5ex>[rr]^f \ar@<-0.5ex>[rr]_g && Y \ar[rr]^q
      \ar[ddrr]_{q'} && Q
      \ar@{.>}[dd]^u \\ \\
& &&& Q' }
\end{displaymath}
\end{definition}


Now we show that in $\Set$, assume

$$R \defeq \Sigma (a_1,a_2 : A) ~ a_1 \sim a_2$$

%$$R = \{(a_1,a_2) : A \times A ~|~ a_1 \sim a_2\}$$

and the two projections are two parallel morphisms $\pi_1,\pi_2 : \morph{R}{A}$,

A quotient corresponds to the coequalizer $(\qset{A},[\_])$:
\begin{displaymath}
    \xymatrix{R \ar@<0.5ex>[rr]^{\pi_1} \ar@<-0.5ex>[rr]_{\pi_2} && A \ar[rr]^{ [\_]}
      \ar[ddrr]_{f} && \qset{A}
      \ar@{.>}[dd]^{\hat{f}} \\ \\
& &&& B}
\end{displaymath}

The factorisation $\hat{\_}$ is just the eliminator, the computation rule and induction principle corresponds to the universal property of it. 

\begin{proposition}
The induction principle implies uniqueness, and is also derivable from the definition of coequalizer.
\end{proposition}

\begin{proof}
It is easy to see that induction principle implies the uniqueness of $\hat{f}$:

Given any $g : \qset{A} \to B$ fulfils the same property as $\hat{f}$, applying induction principle on 

$$\forall (a :A) \to g([a]) = \hat{f}([a])$$

we can deduce that $$\forall (q : \qset{A}) \to g(q) = \hat{f}(q)$$ hence $g = \hat{f}$.

The other way is more difficult:

Given $P : A \to \Prop$, $h : (x : A) \to P (x)$ define

$$P' \defeq \Sigma(x : A) ~P(x) ~\text{and}~ h'(x) = ([x],h(x))$$

we can observe that 

\begin{equation}\label{eq1}
\pi_1 \circ h' = [\_]
\end{equation}

By universal property, there is a unique $\hat{h'}$ s.t. 

\begin{equation}\label{eq2}
\hat{h'} \circ [\_] = h'
\end{equation}

By replacing \ref{eq2} in \ref{eq1}

\begin{equation}\label{eq3}
\pi_1 \circ \hat{h'} \circ [\_] = [\_]
\end{equation}

From \emph{uniqueness} we can easily prove that $[\_]$ is an \emph{epimorphism}.

%as follows:

%given $f_1,f_2 : \qset{A} \to B$ s.t. $f_1 \circ [\_] = f_2 \circ [\_]$

%Because $f_1 \circ [\_] \circ \pi_1 = f_1 \circ [\_] \circ \pi_2$ and $f_2 \circ [\_] \circ \pi_1 = f_2 \circ [\_] \circ \pi_2$ there is a unique $g : \qset{A} \to B$ such that

%$g \circ [\_] = f_1 \circ [\_] = f_2 \circ [\_]$

%Because both $f_1$ and $f_2$ fulfil this property, $f_1 = f_2$, hence $[\_]$ is epi.

Thus from \ref{eq3}, we prove that

$$\pi_1 \circ \hat{h'} = id$$

which implies that any $q : \qset{A}$, the type of $\pi_2(\hat{h'}(q))$ is

$P(\pi_1(\hat{h'}(q))) = P(q)$

as expected, hence we derive the induction principle. 
In fact, following the same procedure, a dependent eliminator is also derivable.
\end{proof}

The coequalizer (quotient) is effective if the following diagram is a \emph{pullback}

\[\xymatrix{
R\pullbackcorner\ar[rr]^{\pi_1}\ar[dd]_{\pi_2} && A\ar[dd]^{[\_]} \\ \\
A\ar[rr]_{[\_]} && Q
}\]

\begin{proof}
Assume we have two points $a, b : \textbf{1} \to A$ satisfying $[ a ] = [ b ]$.

From the pullback property, there is a unique point $r : \textbf{1} \to R$ such that

$$\pi_1(r) = a$$ and $$\pi_2(r) = b$$ 

Hence $(a, b)$ is an element of $R$, by definition it means

$$a \sim b$$
\end{proof}

In \Cref{dq}, we also introduce two other notions: prequotient and definable quotient.

Cateogorically speaking, a \emph{prequotient} is just a \emph{fork}

\[\xymatrix{
R\ar@<0.5ex>[rr]^{\pi_0}\ar@<-0.5ex>[rr]_{\pi_1}& & A\ar[rr]^{[\_]}
& & Q
}\]

and a \emph{definable quotient} corresponds to a \emph{split coequalizer} which is a \emph{fork} with two morphisms $\emb : Q \to A$ and $t : A \to R$ such that 

\begin{itemize}

\item $[\_] \circ \emb = 1_Q$

\item $\emb \circ [\_]  = \pi_2 \circ t$ and 

\item $\pi_1 \circ t = 1_A$

\end{itemize}

In fact, $t$ corresponds to the "complete" property of definable quotients as $t(a) = (\emb [ a ] , a)$.

\section{Quotients as an adjunction}

As Jacobs \cite{Jacobs94quotientsin} suggests, quotient can be described as a left adjoint to an equality functor.

Let us recall the definition first.
\begin{definition}
\textbf{Adjunction}.
Given two categories $A$ $B$, a functor $F : A \to B$ is left adjoint
to $G : B \to A$ if we have a natural isomorphism
$\Phi : \text{hom}_{B}(F ~\_ , \_) \to hom_{A}(\_, G ~\_)$
\end{definition}


Given the category of setoids $\textbf{Setoid}$ and category of sets $\textbf{Set}$, there is an equality functor $\nabla : \textbf{Set} \to \textbf{Setoid}$ defined as
$$\nabla A \defeq (A , =_{A})$$

where the morphism part is trivial embedding.

Quotients can be seen as a functor $\textbf{Q} : \textbf{Setoid} \to \textbf{Set}$ which is
left-adjoint to a equality functor $\nabla A \defeq (A , =_{A})$

The object part of this functor corresponds to the formation rule of quotients, hence we can use $\qset{B}$
to represent $\textbf{Q} ~ (B , \sim)$.

The adjunction can be described by a natural isomorphism

$\Phi : \text{hom}_{\text{Set}}(\textbf{Q}  ~\_ , \_) \to \text{hom}_{\text{Setoid}}(\_ , \nabla ~\_)$

or a diagram for each $(Y, \sim) : \textbf{Setoid}$ and $X : \textbf{Set}$:

\begin{equation*}
\begin{aligned}
\qset{Y} & \to X \\
\midrule
\midrule
(Y, \sim) & \to (X , =_{X})
\end{aligned}
\end{equation*}

%\begin{equation*}
%\begin{aligned}
%\text{hom}_{\text{Sets}}(\qset{Y},X) \\
%\midrule
%\midrule
%\text{hom}_{\text{Setoids}}((Y, \sim), (X , =))
%\end{aligned}
%\end{equation*}


which consists of $\Phi_{(Y, \sim),X}$ and its inverse $\Phi^{-1}_{(Y, \sim),X}$ (the subscripts are omitted later).


Given an identity morphism $\text{id} : \qset{A} \to \qset{A}$, 

$$\Phi(\text{id}) : (A, \sim) \to (\qset{A}, =_{\qset{A}})$$

is just the introduction rule $[\_] : A \to \qset{A}$ with the property that it is compatible with $\sim$.
It is also called \emph{unit} written as $\eta_{(A,\sim)}$.

Given a morphism $f : (A, \sim) \to (B , =_{B})$ which is a function compatible with $\sim$,

$$\Phi^{-1}(f) : \qset{A} \to B$$ 

which corresponds to the elimination rule.

The computation rule $\hat{f} \circ [\_] \equiv f$ corresponds to the following digram in the category of setoids:


\[\xymatrix{
(A,\sim)\ar@<-0.5ex>[dd]_{\eta_{(A,\sim)}}\ar@<0.5ex>[ddrr]^{f} \\ \\
(\qset{A}, =_{\qset{A}})\ar@<-0.5ex>[rr]_{\nabla(\Phi^{-1}(f))} & & (B , =_{B})
}\]

which is commutative because

$\nabla (\Phi^{-1}(f)) \circ \eta_{(A,\sim)}$

$=\Phi(\Phi^{-1}(f))$ by adjunction law $G(f) \circ \eta_{Y} = \Phi(f)$

$=f$



We can also recover the adjunction from the definition of quotients. Define

$$\textbf{Q} ~ (Y , \sim) \defeq \qset{Y}$$

The adjunction is given by

$$\Phi(f) \defeq f \circ [\_] ~\text{and} ~ \Phi^{-1}  (g, g^{\sim}) \defeq \hat{g}$$

The computation rule and induction principle just express that these two mapping are each other inverses.

%which is compatible with $\sim$ because $[\_]$ is compatible with $\sim$.

%We can define its inverse as

%: assume $f : \qset{Y} \to X$, there is a mapping

%It is an isomorphism because it has an inverse: assume $(g, g^{\sim}) : (Y, \sim) \to (X, =_{X})$


%such that

%$\Phi^{-1} (\Phi(f))= \dlift{f \circ [\_]} = f$

%because for all $y : Y$, $\dlift{f \circ [\_]} ([ y ]) \equiv (f \circ [\_]) (y) \equiv f ([ y ])$




%where $g$ is a function compatible with $\sim$

%To prove that this is an isomorphism, assume $f : \qset{B} \to A$, there exists a mapping

%$\Omega ~ f = f \circ [\_]$ and its inverse

%$\Omega^{-1} ~ g = \hat{g}$ (we omit the property that $g$ respects $\sim$)

%The isomorphic properties can be verified as follows,

%$\Omega^{-1} (\Omega~f) = \dlift{f \circ [\_]} = f$ by the uniqueness from lifting.

%$\Omega (\Omega^{-1}~g) = \hat{g} \circ [\_] = g$ by definition of lifting.


\section{Quotients in \hott}\label{qthott}

As we mentioned before, quotient types (in the sense of \ref{iqs}) are available in \hott. Because of the different interpretations of types, it makes less confusion to call them \emph{quotients} or \emph{set quotients} here.

First, let us recall that

\begin{itemize}
\item an h-proposition (hProp) $P$ is a type which has the property $\forall(a, b : A) \to a =_{A} b$, and

\item an hSet $S$ is a type such that forall $x, y : S$, $x =_{S} y$ are h-propositions.
\end{itemize}

For simplicity, we use "sets" for h-sets and "propositions" for h-propositions here. Also $\Prop$ is \emph{not} the built-in universe of propositions in Coq, but the universe of h-propositions.

\subsection{An impredicative encoding of quotient sets}\label{impredicative}

Vladimir Voevodsky has introduced an impredicative definition of quotients which has been encoded in Coq \cite{voe:hset}. 

%\footnote{\url{http://www.cse.chalmers.se/~coquand/cirm.pdf}} 

%Assume we have a type $A$ (which is not necessarily a set) and an equivalence relation $\sim : A \to A \to \Prop$. 


Assume we have a set $A$ and an equivalence relation $\sim : A \to A \to \Prop$. 

\begin{definition}
An \textbf{equivalence class} is a predicate $P : A \to \Prop$ such that

it is inhabited $\exists (a : A) ~P (a)$

For all $x, y : A$,

$P(x) \to P(y) \to x \sim y$

$P(x) \to x \sim y \to P(y) $

The properties can be encoded as 

$$\text{EqClass}(P) \defeq (\exists (a : A) ~ P (a)) \wedge (\forall (x, y : A) \to P(x) \to (x \sim y \iff P(y)))$$
\end{definition}

\begin{definition}
We define the \textbf{set quotient} as

$$\qset{A} \defeq \Sigma (P : A \to \Prop) ~ \text{EqClass}(P)$$
\end{definition}

$\qset{A}$ is a set because $A \to \Prop$ is a set and $\text{EqClass}(P)$ is a proposition. $\wedge$ is the non-dependent $\Sigma$-type for propositions and $\forall$ is the $\Pi$-type for propositions.
Because it is in fact a triple, we use $(P,p,q) : \qset{A}$ to represent an element of it for convenience, 
where $P$ is the predicate, $p$ is the truncated witness of $P$ is inhabited, and $q$ contains the proofs of the logical equivalence.
%where $P : A \to \Prop$, $p : \exists (a : A) ~ P (a)$ and $q : \forall (x, y : A) \to P(x) \to (x \sim y \iff P(y))$.

 The encoding of $\exists(a : A)~ P(a)$ is given by a truncated $\Sigma$-type: $\| \Sigma(a : A)~P(a) \|$.
The (-1)-truncation $\|-\|$ is defined \emph{impredicatively} as

$$\|X\| \defeq \forall (P : \Prop) \to (X \to P) \to P$$

with a trivial embedding function $|\_| : X \to \| X \|$:

$$| x | \defeq \lambda P ~ f \to f(x)$$

We can simply recover the elimination rule for truncation: given any function $f : X \to P$ where $P$ is a proposition, we can define a function of type $\| X \| \to P$
as 

$$\tilde{f}(x) \defeq x(P,f)$$

and $\tilde{f}(|x|) \equiv f(x)$ automatically holds.

\begin{remark}
Note that $\|X\|$ is in the universe of $\Set_1$, but with \textbf{resizing rules} proposed by Voevodsky \cite{Universe-poly,RR}, $\|X\|$ is moved to the universe of $\Set$. We can apply the resizing rule for proposition because $\|X\|$ behaves like a proposition.
It also has to be noticed that it is impossible to extract an element of $A$ from an proof of $\text{EqClass}(P)$ because of the impredicative encoding. 
\end{remark}

There is a canonical function $[\_] : A \to \qset{A}$ corresponding to the \textbf{introduction rule}:

$$[ a ] \defeq (\lambda x \to a \sim x, | a, \text{ref}(a) | , \lambda x~y ~p \to (\lambda q \to\text{trn}(p,q), \lambda q \to \text{trn}(\text{sym}(p),q)))$$

which is compatible with $\sim$. The verification of compatibility requires propositional extensionality and functional extensionality which are available in \hott.
In fact, we can prove that, $[a]$ is an unique representation of an equivalence class.

\begin{lemma}\label{uniquerepresentation}
Given $(P,p,q) : \qset{A}$ and $a : A$ such that $ex : P(a)$ holds, $[a] ={\qset{A}} (P,p,q)$, namely

$$\Sigma(a : A)~P(a) \to [a] ={\qset{A}} (P,p,q)$$ is inhabited.
\end{lemma}
\begin{proof}

Because $\qset{A}$ is a $\Sigma$-type whose second component $\text{EqClass}(P)$ is a proposition depends on the first component, if the first components are equal, i.e.\

$$\lambda b \to a \sim b = P$$

then their second components are also equal because of proof-irrelevance.

By functional extensionality, we only need to prove that

\begin{equation}\label{abpb}
\forall(b : A) \to a \sim b = P(b)
\end{equation}

Recall that the type of $q$ is $\forall(x,y : A) \to P(x) \to (x \sim y \iff (P(y)))$, from assumption $ex : P(a)$, we can prove that 


$$\forall(b : A) \to a \sim b \iff P(b)$$

Then we can simply prove \ref{abpb} by applying propositional extensionality.
Therefore

$$[a] = (P,p,q)$$ 

\end{proof}

A lifting function (non-dependent eliminator) for functions compatible with $\sim$ is also expected. Since we can not extract a element of $A$, it has to be defined in a more complicated way.

\begin{lemma}
Given a function $f : A \to B$ into a \emph{set} $B$ which is compatible with $\sim$, there exists a unique function $\hat{f} = \qset{A} \to B$ such that $\hat{f} ([ a ]) \equiv f(a)$.
\end{lemma}
\begin{proof}
Assume we have an element $(P, p, q) : \qset{A}$,

we can define a function $f_{P} : (\Sigma (x : A) ~P(x)) \to B$ simply by 

$$f_{P} \defeq f \circ \pi_{1}$$

but our witness $p : \| \Sigma (x : A) ~P(x) \|$ is truncated which can not be applied to $f_{P}$. However we can generate a function

$$\bar{f_{P}} : \| (\Sigma (x : A) ~P(x)) \| \to B$$

from \ref{constantT}  if $f_{P}$ is a constant function:

For any two elements of $\Sigma (x : A) ~P(x)$, $(x_1,p_1)$ and $(x_2,p_2)$, by applying one of the property contained in $q$

$$\forall(x,y : A) \to P(x) \to P(y) \to x \sim y$$

to $p_1 : P(x_1)$ and $p_2 : P(x_2)$, we prove that 

$$x_1 \sim x_2$$

Then because $f$ is compatible with $\sim$,

$$f(x_1) = f(x_2)$$

By definition of $f_P$,

$$f_P(x_1,p_1) \equiv f (x_1) = f(x_2) \equiv f_P(x_2,p_2)$$

hence $f_P$ is a constant function.


To summarise, the lifting function can be defined as

$$\hat{f}(P,p,q) \defeq \bar{f_{P}}(p)$$


The \textbf{computational rule} can be verified easily:

$$\hat{f}([a]) \equiv \bar{f}_{\lambda x \to a \sim x}(|a|) \equiv f(a)$$

The \textbf{induction principle} can be generated as follows:

Suppose we have $Q : \qset{A} \to \Prop$, $h : (a : A) \to Q([a])$ and $(P,p,q) : \qset{A}$, we expect the \emph{proposition} $Q(P,p,q)$ holds. Since $p : \| \Sigma (a : A) ~ P(a) \|$, from the elimination rule of truncation, we only need to construct a function of type

$$\Sigma (a : A) ~ P(a) \to Q(P,p,q)$$


Given $(a,ex) : \Sigma (a : A) ~ P(a)$, we know

$$[a] = (P,p,q)$$

from \ref{uniquerepresentation}. Thus we can substitute in $h(a) : Q([a])$ to generate a term of type $Q(P,p,q)$. Therefore we have the induction principle. The uniqueness of $\hat{f}$ is simply implied by the induction principle.
\end{proof}


The following lemma is suggested by Nicolai Kraus and can be found in \cite{krausEscardoEtAll_existence}.

\begin{lemma}\label{constantT}
Given a constant function $g : X \to Y$ where $Y$ is a set, hence it satisfies

$$\forall(x,y : X) \to g(x) = g(y)$$

there exists a function $\bar{g} : \|X\| \to Y$ such that $\bar{g}(| x |) \equiv g(x)$.
\end{lemma}

\begin{proof}

Define the subset

$$Y' \defeq \Sigma (y : Y)~ \|\Sigma (x : X) ~g(x) = y \|$$

Intuitively it contains only the image of the constant function i.e.\ \textbf{$Y'$
is propositional}:

For any $(y_1,p_1) : Y'$ and $(y_2,p_2): Y'$, 

we can first generate the proofs 

$p_1 ((g(x) = y_1),\pi_2) : g(x) = y_1$ and 

$p_2 ((g(x) = y_2),\pi_2) : g(x) = y_2$

By symmetry and transitivity we can prove that $y_1 = y_2$.

From the fact that truncated type is always propositional, we can also deduce that $p1 = p2$, then $(y_1,p_1) = (y_2,p_2)$. Hence we can conclude that $Y'$ is propositional.

We can simply define a function $g' : X \to Y'$ from $g$ as

$$g'(x) \defeq (g(x), \lambda Q ~f \to f (x , \text{refl}_{=}(g(x))))$$


Because $Y'$ is propositional, it is possible to lift $g'$ to a function $\tilde{g'} : \| X\| \to Y'$ which is technically defined as

$$\tilde{g'}(x) \defeq x (Y',g')$$

Finally we define

$$\bar{g} \defeq \pi_1 \circ \tilde{g'}$$

Which fulfils the computation rule

$$\bar{g}(|x|) \equiv  \pi_1 (|x|(Y',g'))  \equiv \pi_1 (g' (x)) \equiv g(x)$$
\end{proof}


Furthermore, since propositional extensionality is a special case of univalence, we can prove that the impredicative quotients are effective from \Cref{PUEF}.

\begin{theorem}
In \hott, the impredicative encoding of quotient sets gives rise to all the rules of quotients in the sense of \ref{iqs} including effectiveness.
\end{theorem}

\subsection{Quotient inductive types}


%As we mentioned before, quotient types (in the sense of \ref{iqs}) are available in \hott. Because of the different interpretations of types, it makes less confusion to call them \emph{quotients} here.

An alternative way to define quotients in \hott is by using higher inductive types.


Assume $A$ is a set and $\_\sim\_ : A \to A \to \Prop$ is an equivalence relation. 
To build a quotient, we can simply impose level-$1$ morphisms in the structure of given sets according to the equivalence relation.
Thus, a quotient $\qset{A}$ can be defined as a higher inductive type:

\begin{itemize}
\item $[\_] : A \rightarrow \qset{A}$
\item $eqv : (a,b : A) \rightarrow a \sim b \rightarrow  [a] = [b]$
\item $isSet : (x,y:\qset{A}) \rightarrow (p_1,p_2 : x = y) \rightarrow p_1 = p_2$
\end{itemize}

It is also a set so we call it \textbf{set-quotient} or \textbf{quotient inductive types} (QITs).

\begin{remark}
Categorically, there is no difference between a set and a setoid because a skeleton of a setoid
is just a discrete category which is equivalent to a set.
\end{remark}

Some examples suggest that QITs are more powerful than quotient types.

One example is the definition of real numbers $\R$ which will be discussed in \Cref{rl}. Our construction of reals by Cauchy sequences of rational numbers is not Cauchy complete because not all  equivalence classes have a limit. However, the Cauchy approximation approach (see Subsection 11.3.1 in \cite{hott}) using quotient inductive types is Cauchy complete due to the fact that the equivalence relation and limits are included in its definition.

Another example is unordered trees (rooted tree) which are trees connected to a multiset of subtrees, hence there is no ordering on subtrees.

Firstly we define ordered trees as:

\begin{itemize}
\item A leaf $l: \mathsf{Tree}$, or
\item An ordered list of subtrees indexed by $I$, $st : (I \rightarrow \mathsf{Tree}) \rightarrow \mathsf{Tree}$,
\end{itemize}

With the following equivalence relation:

\begin{itemize}
\item $l_{eq} : l \sim l$,
\item $st_{eq} : (f , g : I \to \mathsf{Tree}) \to f \sim_{p} g \to sp(f) \sim sp(g)$
\end{itemize}

where $f \sim_{p} g$ stands for $f$ is a permutation of $g$. The permutation can be defined using a bijective map $p : I \to I$ which relates equivalent subtrees recursively.

%However, if we use quotient type $\mathsf{Tree^{\sim}} := \qset{\mathsf{Tree}}$, the resulting trees have unordered subtrees which are themselves ordered.

If we define unordered trees as a quotient type $\mathsf{Tree^{\sim}} := \qset{\mathsf{Tree}}$, 
it is problematic to lift the constructor $st$, i.e.\ to define $\hat{st}$.
For trees with finite subtrees such as \emph{binary trees} where $I \defeq \bool$, it can be lifted by nesting lifting functions,

$$\hat{st}(a,b) = \dlift{\hat{st}(a)}(b)$$

because its type is isomorphic to $\mathsf{BTree} \rightarrow \mathsf{BTree} \rightarrow \mathsf{BTree}$.
Intuitively this approach can be applied to trees with finite subtrees. However it fails if have infinite subtrees, for example when $I \defeq \N$.

However if we use QITs to define unordered trees, we can define the equivalence relation simultaneously with constructors for types so that:

\begin{itemize}
\item $l: \mathsf{Tree}$, 
\item $st : (I \rightarrow \mathsf{Tree}) \rightarrow \mathsf{Tree}$

and a set of paths relates two permuted trees:

\item $l_{eq} : l  =_{\mathsf{Tree}} l $

\item$st_{eq} : \forall (f, g : I \to \mathsf{Tree}) \rightarrow  f \sim_{p} g \rightarrow  st(f) =_{\mathsf{Tree}} st(g)$
\end{itemize}

Thus we avoid the problems of lifting $st$ because the equivalence relation has become the internal equality of this type.


Similarly the cumulative hierarchy of all sets introduced in \cite{hott} (see section 10.5) also suggests that quotient types have some weaknesses compared to quotient inductive types.

A cumulative hierarchy can be given by constructors,

$$\{\_\} : (I : \Set) \to (I \to M_0) \to M_0$$

along with a subset relation,

$$\_\in\_ : M_0 \to M_0 \to \Prop$$

which is inhabited if $ f(i) \in \{ I , f \}$.

Then we can easily define the equivalence relation on "sets" using set-theoretical definition

$A \sim B \defeq \forall m : M_0, m \in A \iff m \in B$

Similar to unordered trees, we can not obtain the constructor $\dlift{\{\_\}}$ because the index set $I$ can be infinite.

To summarise, it seems that quotient inductive types are more powerful than quotient types due to the ability of defining term constructors and equivalence relations simultaneously. 
However, quotient inductive types are not available in type theories other than \hott and the computational interpretation of them is still an open problem.
Moreover, there can be more general quotients in \hott, for example a quotient of a type by a $1$-groupoid (See section 9.9 in \cite{hott}). It is interesting to investigate \emph{real} \emph{quotient types} in \hott, but it is beyond this thesis.

\section{Related work}\label{related}

The introduction of quotient types in Type Theory has been studied by several authors in different versions of \mltt and using various approaches.

\begin{itemize}

\item In \cite{cab}, Mendler et al. have first considered building new types from a
given type using a quotient operator $//$. Their work is done in an
implementation of \ett, NuPRL. 

In NuPRL, given the base type $A$ and the an equivalence relation $E$, the quotient is denoted as $(x,y) : A//E$. Since every type comes with its own equality relation in NuPRL, the quotient operator can be
seen as a way of redefining equality for a type.

They also discuss problems that arise from defining functions on the new type which can be illustrated using a simple example:

When we want to define a function $f : (x,y) : A//E \to \bool$, it is in fact defining a function on $A$. Assume $a, b : A$ such that $E(a,b)$ but $f(a) \not= f(b)$. This will lead to
inconsistency since $E(a,b)$ implies $a$ converts to $b$ in \ett, hence
the left hand side $f(a)$ can be converted to $f(b)$, namely we get $f(b) \not= f(b)$
which is contradicted with the equality reflection rule. 

Therefore a function is well-defined \cite{cab} on the new type only
if it respects the equivalence relation $E$, namely

$$\forall (a,b : A) \to E(a,b) \to f(a) = f(b)$$


 After the introduction of quotient types, Mendler further investigates
 this topic from a categorical perspective in ~\cite{men:90}. He uses
 the correspondence between quotient types in \mltt{} and coequalizers
 in a category of types to define a notion called \emph{squash types},
 which is further discussed by Nogin \cite{nog:02}.

\item Nogin \cite{nog:02} considers a modular approach to axiomatizing quotient types in NuPRL as well. He discusses some problems about quotient types. For example, since the equality is
extensional, we cannot recover the
witness of the equality.  He suggests including more axioms to
conceptualise quotients. He decomposes the formalisation of quotient type
into several smaller primitives such that they can be handled much
simpler.


\item Jacobs \cite{Jacobs94quotientsin} introduces a syntax for quotient types based on a predicate logic
within simple type theory. He discusses quotient types from a categorical theoretical perspective. In fact the syntax of quotient types arises from an adjunction as we mentioned before.

\item To add quotient types to \mltt, Hofmann proposes three models for
quotient types in \cite{hof:phd}. The first one is a setoid model for
quotient types. In this model all types are attached with partial
equivalence relations, namely all types are partial setoids rather than
sets. It does not provide dependency at the level of types but only at the level of the relations. 
%This is similar to \ett in some sense but retains computational property in \itt. 
The second one is groupoid model which supports most features required 
but it is not definable in \itt. 
He also proposes a third model to as an attempt to overcome problems in the previous two models. More type dependency is provided and quotient types are believed to be definable in this model.
However it also has some disadvantages. 
he also shows that \ett is conservative over \itt extended with quotient types \cite{hof:95:con}.


\item Altenkirch \cite{alti:lics99} also provides a different setoid model which is built in an \itt
extended with proof-irrelevant universes of propositions and $\eta$-rules for $\Pi$-types and $\Sigma$-types. It is decidable, adequate and permits large eliminations. We have implemented this setoid model and interpret quotient types in it (see \Cref{models}).

\item Homeier \cite{hom} also axiomatises quotient types in Higher Order Logic
(HOL), which is also a theorem prover. He creates a tool package to
construct quotient types as a conservative extension of HOL such that
users are able to define new types in HOL. Next he defines the
normalisation functions and proves several properties of
these. Finally he discussed the issues when quotienting on the
aggregate types such as lists and pairs.


\item Courtieu \cite{cou:01} shows an extension of Calculus of Inductive Constructions
with \emph{Normalised Types} which are similar to quotient types, but equivalence relations are replaced by normalisation functions which select a canonical element for each equivalence class. 
In fact normalised types can be seen as a proper subset of quotient types. We can easily recover a quotient type from a normalised type as below

$$a \sim b \defeq [ a ] = [ b ]$$

However not all quotient types have normal forms, for example, the set of real numbers (see \Cref{rl}).
The notion \emph{definable quotients} we proposed in \Cref{dq} is also similar to it, but does not provide a new type automatically.

\item Barthe and Geuvers \cite{bar:96} also propose a new notion called
\emph{congruence types}, which is also a special class of quotient
types, in which the base type are inductively defined and with a set
of reduction rules called the term-rewriting system. The idea behind
it is the $\beta$-equivalence is replaced by a set of
$\beta$-conversion rules. Congruence types can be treated as an
alternative to the pattern matching introduced in \cite{coq:92}. The main
purpose of introducing congruence types is to solve problems in
term rewriting systems rather than to implement quotient types.
Congruence types are not inductive but have good computational behaviour because we can use term-rewriting system to link a term of base type to a unique term of congruence type as its normal form.
However this approach has some problems in termination criteria and interaction between rewriting systems \cite{cou:01}.

\item Barthe and Capretta \cite{bar:03} compare different ways to setoids in Type Theory.
The setoid is classified as partial setoid or total setoid depending
on whether the equality relation is reflexive or not. They also
consider obtaining quotients with different kinds of setoids, especially
the ones from partial setoids.
In their framework of partial setoids, suppose we have a partial setoid $(A,\sim)$,  an element $x : A$ such that $x \sim x$ is called \emph{defined} elements, the others are undefined. Therefore if we simply define a quotient by replacing underlying partial equivalence relation with a new one $R$, undefined elements in the base setoid may be incorrectly introduced in the quotient. The reason is that there possibly exist some undefined elements $x : A$ satisfying $R(x,y)$. They solve the problem by defining a restricted version of $R$ which only relates defined elements.


\item Abbott, Altenkirch et al. \cite{abb:04} provides the basis for
programming with quotient datatypes polymorphically based on their
works on containers which are datatypes whose instances are
collections of objects, such as arrays, trees and so on. Generalising
the notion of container, they define quotient containers as the
containers quotiented by a collection of isomorphisms on the positions
within the containers.

\item Voevodsky \cite{voe:hset} implements quotients in Coq based on a set
of axioms of \hott. He firstly implement
equivalence class and use it to implement quotients which is an
analogy to the construction of quotient sets in set theory. The details has been given in \Cref{impredicative}.
\end{itemize}





\section{Summary}

We have given the syntax of quotient types in this chapter. The relation is required to be an equivalence in our definition which is different to \cite{hof:95:sm}. In fact, the equivalence condition does not affect the construction of quotient types. Jacobs  \cite{Jacobs94quotientsin}  has shown that, for arbitrary relation $R$, the same construction can be interpreted as set theoretical quotient sets of $A/R^{\equiv}$, where $R^{\equiv}$ is the equivalence closure of $R$. However the equivalent condition does allow us to prove that 

 Two approaches of defining elimination rules has been given, one is a combination of non-dependent eliminator with an induction principle as in Hofmann's definition and another is a dependent eliminator. We have also shown they are equivalent. 

We have also shown that propositional extensionality actually implies the effectiveness of quotients. The characterisation of quotients in category theory has been given. It does not only corresponds to coequalizers but also can be generated from a left adjoint functor to the equality functor $\nabla : \textbf{Set} \to \textbf{Setoid}$. We have also briefly reviewed the research of quotient types by other authors.






