\acknowledgements{\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics
%\begin{acknowledgements}

The first person I would like to thank is my supervisor Thorsten Altenkirch. He offered me a great internship opportunity about encoding numbers in Agda, which was later developed into this project. He has always been patient in explaining my questions and taught me a lot on research. I would also like to thank him for his guidance and feedbacks on this thesis.
I would also thank my second supervisor Thomas Anberrée who introduced me functional programming and he has given me a lot precious advices on my project.
I would also like to thank Venanzio Capretta who examined my first and second year reports and provided me helpful feedbacks.

I would like to thank Ambrus Kaposi and Nicolai Kraus for their great help in my thesis writing and comments on my drafts.
My friends Nicolai Kraus, Ambrus Kaposi, Christian Sattler, Paulo Caprotti, Florent Balestrieri, Gabe Dijkstra, Ivan Perez, Neil Sculthorpe and all other PhD students in our lab has helped me a lot in my project by either teaching me mathematics, discussing on research topics and playing badminton to exercise our bodies. 
The Functional programming lab is like a lively family and I really enjoy the time here. I would like to thank everyone here, without their kind help, the thesis would not have been finished.


I would also like to thank the organizers and
other participants of the special year on \hott at the
Institute for Advanced Study where we had many interesting
discussion topics some of which are related to part of this work, especially Guillaume
Brunerie whose proposal made it possible. I would also thank Nordvall Forsberg who provided important discussions on our work.

I would also like to thank my parents and other family members.
During the years in Nottingham, My mother Guangfei Lv, and my father Youyuan Li have always been very supportive of me, even though I am living far from them. They talked to me a lot and sent my favourite foods to me many times. My aunt Guangshu Lv who is living in Germany also helps me a lot. She has given my many  advices and has sent mails to me. Without their support I would never achieve my goals.

Finally, I would like to thank the School of Computer Science and international office in the University of Nottingham who financially support this project by providing important studentship to me.	

%\end{acknowledgements}

}